\vspace{-1in} Demystifying the Computer Science \\PhD Admission in US Universities\\{\large A Guide for International Students

### 2023-11-06 demystify.tex
updates

One main concern for non-CS or non-STEM students is if you have the sufficient technical background, typically obtained through core CS courses.  So you will want to show that you have such knowledge through your coursework, projects, or research. 
For example, if you have taken a course on Algorithms, even online ones like Coursera, you can talk about it in your statement.  If you have done a project or have a professional certification (e.g., A+) through work that requires knowledge of OS, you can talk about it in your statement.  If you have done research that requires knowledge of Discrete Maths, you can talk about it in your statement.  You can also ask your LoR writers to talk about your technical background.  
In summary, in your application, convince us that you have the technical background to do CS PhD research. 
Note that many online courses teaching about AI does not really help much as these are not core CS knowledge.

In short, you \emph{do not need} to formally taking CS courses, you just need to show that you have these essential knowledge through ways as mentioned. Many universities are well-aware that incoming graduate students might not have all the technical background, so they often have a "bridge" courses to help students catch up.  For example, GMU has 4 bridge courses (Data Structure and Algorithm, Computer Systems, Discrete Maths, and Programming Foundation) that incoming students can take to catch up on their CS knowledge.

\begin{commentbox}[Vu:]
    I would strongly advocate to admit a student with a non-STEM background (e.g., business, English major) but can show that they have a strong drive for CS by studying core CS knowledge through various channels (e.g., self-study through online courses, projects, etc).  I have seen many students with non-CS background who are very successful in CS PhD.  In fact, I have seen many students with CS background who are not successful in CS PhD.  So it is not about your background, it is about your drive and passion for CS research.    \end{commentbox}

@@

\paragraph{About Me} I am an assistant professor at George Mason University (GMU). Prior to GMU, I was at the University of Nebraska-Lincoln (UNL). I have been involved in the PhD admission process at GMU and UNL for many years and currently I am the program director of the Masters program in Software Engineering at GMU (thus also have some experience with the MS admission process--which is quite different than PhD). I got my undergraduate and MS degrees from Penn State. I did my PhD at the University of New Mexico-Albuquerque, and then a postdoc at the University of Maryland-College Park.  I had also worked at the Naval Research Lab and Lockheed Martin.

While I grew up in the US, I am familiar with international students, e.g., many of my students are international students.  I have many international collaborators and also have been a mentor for many students from Vietnam.  I also have many colleague and friends who were international students and have heard their stories.  I hope to capture their challenges and experiences in this document so that it can be useful for prospective students.

>>>>

One main concern for non-CS or non-STEM students is if you have the sufficient technical background, typically obtained through core CS courses.  So you will want to show that you have such knowledge through your coursework, projects, or research. 
For example, if you have taken a course on Algorithms, even online ones like Coursera, you can talk about it in your statement.  If you have done a project that requires knowledge of OS or have a professional certification (e.g., A+) through work, you can talk about it in your statement.  If you have done research that requires knowledge of Discrete Maths, you can talk about it in your statement.  You can also ask your LoR writers to talk about your technical background.  
In summary, in your application, convince us that you have the technical background to do CS PhD research. 
Note that many online courses teaching about AI does not really help much as these are not core CS knowledge.

In short, you \emph{do not need} to formally taking CS courses, you just need to show that you have these essential knowledge through ways as mentioned. Many universities are well-aware that incoming graduate students might not have all the technical background, so they often have a "bridge" courses to help students catch up.  For example, GMU has four bridge courses (Data Structures and Algorithms, Computer Systems, Discrete Math, and Programming Foundation) that incoming students can take to catch up on their CS knowledge.

\begin{commentbox}[Vu:]
    I would strongly advocate for a non-STEM student who shows that they have a strong drive for CS by studying core CS knowledge through various channels (e.g., self-study through online courses, projects, etc).  I have seen many students with non-CS background who are very successful in CS PhD.  I also have seen many students with CS background who are not successful in CS PhD.  So it is not about your background, it is about your drive and passion for CS research.
\end{commentbox}

@@

\paragraph{About Me} I am an assistant professor in the CS dept at George Mason University (GMU). Prior to GMU, I was at the University of Nebraska-Lincoln (UNL). I have served in the PhD admission process at GMU and UNL for many years.  Currently I am also the program director of the Masters program in Software Engineering at GMU (thus also have some experience with the MS admission process--which is quite different than PhD). I got my undergraduate and MS degrees from Penn State. I completed my PhD at the University of New Mexico-Albuquerque and a postdoc at the University of Maryland-College Park.

While I grew up in the US, I am familiar with international students, e.g., many of my students are international students.  I have many collaborators from various countries and have been a mentor for many students from Vietnam.  I also have many colleague and friends who were international students and have heard their stories.  I hope to capture their challenges and experiences in this document so that it can be useful for prospective international students.
